Switch equivalence and weak prime domains for fusions

### 2024-03-27 icalp/Rebuttal.txt

We thank the reviewer for carefully reading the paper and for their comments. Here are some answers:

REVIEWER 1
We believe that the paper fits into "Algebraic and Categorical Models of Computation" which is one of the topics of ICALP. We did our best to make the paper accessible to people having a basic knowledge of category theory and we inserted a number of examples aimed at communicating the main problems and solution of dealing with non-linear rewriting (which apply to a number of cincrete visual rewriting formalisms like string diagrams [9]). In the preparation of the final version we can put further effort in conveying intuition.\

REVIEWER 2
- "it is unfortunate that you did not find space for describing the class in any detail, but please at least indicate how it compare to the class studied in [6]."

As mentioned in the paper, the class is the natural generalisation of the one in [6] from adhesive to M-adhesive categories. In the final version we can make this more explicit and add some further details.

- Ex. 3.7 does not look quite right. The was not adhesive, wasn't it? It was only iso-adhesive.

The idea here is to provide an example of a M-left linear rewriting systems over an M-adhesive category which is not strong enforcing. Hence left-linear and adheive should read M-left linear and M-adhesive. Omitting the M- prefix which is done also in other parts of the paper here is definitively confusing and we will amend this.

REVIEWER 3

- Is Church-Rosser essentially a consequence of the definition of strong independence pair?

Yes, the idea here was to distill the minimal assumptions needed to make Church-Rosser result work. This is then used to show that for a large class of M-adhesive rewriting systems (the one generalising [6]) this assumptions reduce to the existence of an independence pair.

- well-switching rewriting systems feel overly restrictive and, as such, the development feels slightly preliminary.

Actually, the counterexamples discussed in the paper (ex. 3.14) suggest that out of the class of well-switching rewriting systems no reasonable theory of rewriting can be developed. And, as the reviewer correctly observe, despite the fact that the condition could look restrictive, classes of rewriting systems (Node-merging graphical rewriting systems and E-graphs) which cover many concrete applications in the literature are well-switching.

### 2024-03-27 icalp/Rebuttal.txt

We thank the reviewers for the careful reading of our paper and for their comments. Here are some answers:

REVIEWER 1
We believe that the paper fits into "Algebraic and Categorical Models of Computation" which is one of the topics of ICALP. We did our best to make the paper accessible to people having a basic knowledge of category theory and we inserted a number of examples aimed at communicating the main problems and solution of dealing with non-linear rewriting (which apply to a number of concrete visual rewriting formalisms like string diagrams [9]). In the preparation of the final version we can put further effort in conveying intuition.

@@

REVIEWER 2
- "it is unfortunate that you did not find space for describing the class in any detail, but please at least indicate how it compare to the class studied in [6]."

Indeed, the class is the natural generalisation of the one in [6] from adhesive to M-adhesive categories. In the final version we can make this more explicit and add some further details.

@@

- Ex. 3.7 does not look quite right. The was not adhesive, wasn't it? It was only iso-adhesive.

The idea here is to provide an example of a M-left linear rewriting systems over an M-adhesive category which is not strong enforcing. Hence left-linear and adhesive should read M-left linear and M-adhesive. Omitting the M- prefix here is definitively confusing and we will amend this.

@@

REVIEWER 3

- Is Church-Rosser essentially a consequence of the definition of strong independence pair?

Yes, the idea here was to distill the minimal assumptions needed to make Church-Rosser work. This is then used to show that for a large class of M-adhesive rewriting systems (the one generalising [6]) this assumption reduce to the existence of an independence pair.

@@

- well-switching rewriting systems feel overly restrictive and, as such, the development feels slightly preliminary.

Actually, the counterexamples discussed in the paper (ex. 3.14) suggest that out of the class of well-switching rewriting systems no reasonable theory of rewriting can be developed. And, as the reviewer correctly observes, despite the fact that the condition could look restrictive, classes of rewriting systems (Node-merging graphical rewriting systems and E-graphs) that cover many concrete applications in the literature are well-switching.

### 2024-03-27 icalp/Rebuttal.txt

We thank the reviewers for the careful reading of our paper and for their comments. Here are some answers:

REVIEWER 1
We believe that the paper fits into "Algebraic and Categorical Models of Computation" which is one of the topics of ICALP. We did our best to make the paper accessible to people having a basic knowledge of category theory and we inserted a number of examples aimed at communicating the main problems and solution of dealing with non-linear rewriting (which apply to a number of concrete visual rewriting formalisms like string diagrams [9]). In the preparation of the final version we can put further effort in conveying intuition.

REVIEWER 2
- "it is unfortunate that you did not find space for describing the class in any detail, but please at least indicate how it compare to the class studied in [6]."

Indeed, the class is constructed using a generalisation of the methods exploited in [6], adapting them to the more general M-adhesive context. In the final version we can make this more explicit and add some further details.

@@

- Ex. 3.7 does not look quite right. The was not adhesive, wasn't it? It was only iso-adhesive.

The idea here is to provide an example of a M-left linear rewriting systems over an M-adhesive category which is not strong enforcing. Hence left-linear and adhesive should read M-left linear and M-adhesive. Omitting the M- prefix which is done also in other parts of the paper here is definitively confusing and we will amend this. It is worth noticing that, having in mind the results of [6] and of our Appendix B, in every category built from Set using the comma construction or taking the category of functors is strong enforcing, thus every example of an independence pair which is not strong, must live in a rather "pathological" category.

### 2024-03-28 icalp/Rebuttal.txt

REVIEWER 2
- "it is unfortunate that you did not find space for describing the class in any detail, but please at least indicate how it compare to the class studied in [6]."

Indeed, the class is constructed using a generalisation of the methods exploited in [6], adapting them to the more general M-adhesive context. In the final version we can make this more explicit and add some further details.

- Ex. 3.7 does not look quite right. The was not adhesive, wasn't it? It was only iso-adhesive.

The idea here is to provide an example of a M-left linear rewriting systems over an M-adhesive category which is not strong enforcing. Hence left-linear and adhesive should read M-left linear and M-adhesive. Omitting the M- prefix which is done also in other parts of the paper here is definitively confusing and we will amend this. It is worth noticing that, having in mind the results of [6] and of our Appendix B, in every category built from Set using the comma construction or taking the category of functors is strong enforcing, thus every example of an independence pair which is not strong, must live in a rather "pathological" example of M-adhesive category.

### 2024-03-28 icalp/Rebuttal.txt

We thank the reviewers for the careful reading of our paper and for their comments. Here are some answers:

REVIEWER 1
We believe that the paper fits into "Algebraic and Categorical Models of Computation" which is one of the topics of ICALP. We did our best to make the paper accessible to people having a basic knowledge of category theory and we inserted a number of examples aimed at communicating the main problems and solutions of dealing with non-linear rewriting (which apply to a number of concrete visual rewriting formalisms like string diagrams [9]). In the preparation of the final version we can put further effort in conveying intuition.

@@

The idea here is to provide an example of a M-left linear rewriting systems over an M-adhesive category which is not strong enforcing. Hence left-linear and adhesive should read M-left linear and M-adhesive. Omitting the M- prefix which is done also in other parts of the paper here is definitively confusing and we will amend this. It is worth noticing that, having in mind the results of [6] and of our Appendix B, in every category built from Set using the comma construction or taking the category of functors is strong enforcing, thus every example of an independence pair which is not strong, must live in a rather "pathological" example of M-adhesive category.

REVIEWER 3

- Is Church-Rosser essentially a consequence of the definition of strong independence pair?

Yes, the idea here was to distill the minimal assumptions needed to make Church-Rosser work. This is then used to show that for a large class of M-adhesive rewriting systems (the one generalising [6]) this assumption reduces to the existence of an independence pair.

### 2024-03-28 icalp/Rebuttal.txt

We thank the reviewers for the careful reading of our paper and for their comments. Here are some answers:

REVIEWER 1
- Adequate for ICALP?

We believe that the paper fits into "Algebraic and Categorical Models of Computation" which is one of the topics of ICALP. We did our best to make the paper accessible to people having a basic knowledge of category theory and we inserted a number of examples aimed at communicating the main problems and solutions of dealing with non-linear rewriting (which apply to a number of concrete visual rewriting formalisms like string diagrams [9]). In the preparation of the final version we can put further effort in conveying intuition.

REVIEWER 2
- It is unfortunate that you did not find space for describing the class in any detail, but please at least indicate how it compare to the class studied in [6].

@@

REVIEWER 3

- Is Church-Rosser essentially a consequence of the definition of strong independence pair?

Yes, the idea here was to distill the minimal assumptions needed to make Church-Rosser work. This is then used to show that for a large class of M-adhesive rewriting systems (the one generalising [6]) this assumption reduces to the existence of an independence pair.

- Well-switching rewriting systems feel overly restrictive and, as such, the development feels slightly preliminary.

### 2024-04-13 fusione.tex
quozienti

@@

NOTES ON THE FUNDAMENTAL LEMMA WE NEED

@@

@@

\section{A note on fillers and sequential independence}\label{app:fill}
 In \Cref{prop:equi} we proved that, in the linear case, the existence of an indpendence pair between two derivation is equivalent to that of a filler between them. This result can be further refined: in a\cite{baldan2011adhesivity} a class  $\mathbb{B}$ of (quasi)adhesive category is defined for which the local Church-Rosser Theorem holds even for left-linear DPO-rewriting system. In our language, and given \Cref{prop:fil} and \Cref{rem:locCR}, this amount to prove that, for elements of $\mathbb{B}$, every independence pair induces a filler.

>>>>

\subsubsection{Final pullback complements}

\subsubsection{Factorization systems and $\mathcal{M}$-adhesivity} \todo{Da sistemare}
 Let us start by recalling the notion of a factorization system \cite{bousfield1977constructions,kelly1991note,rosicky2007factorization,tholen1983factorizations}.

\begin{definition}\label{def:fs}\index{factorization system} \index{left lifting property} \index{factorization system!proper -} \index{factorization system!stable -}
	Let $\X$ be a category and $\mathcal{E}$, $\mathcal{M}$ two classes of arrows, we will say that $(\mathcal{E},\mathcal{M})$ is a (orthogonal) \emph{factorization system} if:
	\begin{enumerate}
		\item every isomorphism is in both $\mathcal{E}$ and $\mathcal{M}$;
		\item $\mathcal{E}$ and $\mathcal{M}$ are closed under composition;

		\item  every arrow $f\colon X\to Y$ of $\X$ admits a \emph{$(\mathcal{E}, \mathcal{M})$-factorization}, i.e.~there are arrows $e_f\in \mathcal{E}$ and $m_f\in \mathcal{M}$ with the property that $f=m_f\circ e_f$;
		\item every $e\in \mathcal{E}$ has the \emph{left lifting property} with respect to every $m\in \mathcal{M}$: for every commutative square
		\[\xymatrix{X \ar[r]^g \ar[d]_{e}& Z \ar[d]^{m}\\ Y \ar[r]_{f} \ar@{.>}[ur]^{k}& V}\] 	with $e\in \mathcal{E}$ and $m\in \mathcal{M}$ there exists a unique $k\colon Y\rightarrow Z$ such that 
		\[m\circ k=f \qquad k\circ e=g\]
	\end{enumerate}

	A factorization system is \emph{proper} if every $e\in \mathcal{E}$ is epi and every $m\in \mathcal{M}$ is mono; it's \emph{stable} if for every pullback square ast the one below,  $e\in \mathcal{E}$ implies $e'\in \mathcal{E}$.
	\[\xymatrix{P\ar[r]^{g} \ar[d]_{e'} & X \ar[d]^{e} \\ Z \ar[r]_{f}  & Y}\]
\end{definition}

The following proposition assures us that the factorization of an arrow is unique up to isomorphism.

\begin{proposition}\label{prop:iso}
	Let $(\mathcal{E}, \mathcal{M})$ be a factorization system on a category $\X$. If $e\colon X\to Y$, $e'\colon X\to Y'$ and $m\colon Y\to Z$, $m'\colon Y'\to Z $ are arrows, respectively, in $\mathcal{E}$ and $\mathcal{M}$ such that
	$e'\circ m'=e\circ m$,  then  there exist a unique isomorphism $\phi\colon Y \rightarrow Y$ such that the following diagram commutes
	\[\xymatrix@C=35pt@R=20pt{X \ar[r]^{e'} \ar[d]_{e}& Y' \ar[d]^{m'}\\ Y \ar[r]_{m} \ar@{.>}[ur]^{\phi}& V}\]
\end{proposition}
\begin{proof}Using the left lifting property we get two commutative diagrams:
	\[\xymatrix@C=35pt@R=20pt{
		X\ar[r]^{e'}  \ar[d]_{e}& Y'  \ar[d]^{m'} &  X 	\ar[d]_{e'} \ar[r]^{e} &Y \ar[d]^{m}\\
		Y \ar[r]_{m} \ar[ur]^{f} & Z & Y'  \ar[r]_{m'}\ar[ur]^{g} & Z
	}\]
	thus
	\[	\begin{split}
		m'\circ f \circ g &= m\circ g \\&= m'
	\end{split} \qquad 
	\begin{split}
		f \circ g \circ e' &= f\circ e \\&= e'
	\end{split}\qquad 
	\begin{split}
		m\circ g \circ f &= m'\circ f \\&= m
	\end{split} \qquad 
	\begin{split}
		g \circ f \circ e &= g\circ e' \\&= e
	\end{split}
	\]
	So we have two square
	\[\xymatrix@C=35pt@R=20pt{
		X\ar[r]^{e}  \ar[d]_{e}& Y  \ar[d]^{m} &  X 	\ar[d]_{e'} \ar[r]^{e'} &Y '\ar[d]^{m'}\\
		Y \ar[r]_{m} \ar[ur]^{g\circ f} & Z & Y'  \ar[r]_{m'}\ar[ur]^{f\circ g} & Z
	}\]		
	and the thesis follows from the uniqueness half of the left lifting property.
\end{proof}

\begin{corollary}\label{cor:iso} Given a factorization system $(\mathcal{E}, \mathcal{M})$ on a category $\X$, the following hold:
	\begin{enumerate}
		\item an arrow $f\colon X\to Y$ is in $\mathcal{E}$ (in $\mathcal{M}$) if and only if $m_f$ ($e_f$) is an isomorphism;
		\item $f\in \mathcal{E}$ and $f\in \mathcal{M}$ if and only if $f$ is an isomorphism;
		\item  if $(\mathcal{E}, \mathcal{M})$ is proper, then $g\circ f$ is in $\mathcal{M}$ (in $\mathcal{E}$) implies $f\in \mathcal{M}$ ($g\in \mathcal{E}$).
	\end{enumerate}
\end{corollary}
\begin{proof}
	\begin{enumerate}
		\item  ($\Rightarrow$) By hypothesis $f=\id{Y}\circ f$ ($f=f\circ \id{X}$) is a factorization with $\id{Y}\in \mathcal{M}$ and $f\in \mathcal{E}$ ($\id{X}\in \mathcal{E}, f\in \mathcal{M}$), so the thesis follows from \cref{prop:iso}.

		\smallskip \noindent
		($\Leftarrow$) $f=m_f\circ e_f$, thus if $m_f$ ($e_f$) is an isomorphism then we have $f$ is the composition of two arrows in $\mathcal{E}$ ($\mathcal{M}$) and we can conclude.

		\item This follows immediately from the previous point.

		\item 
		Factor $f$ and $g$ as $m_f\circ e_f$ and $m_g\circ e_g$, let also $h$ be $e_g\circ m_f$ and factor it as $m_h\circ e_h$ so that we get 
		\[\xymatrix@C=45pt{& C \ar[dr]^{m_h}&\\
			A\ar[ur]^{e_h} \ar[rr]^{h}  \ar[dr]^{m_f}& &B \ar[d]^{m_g}\\
			X \ar[u]^{e_f} \ar[r]_f& Y \ar[r]_g \ar[ur]^{e_g} & Z}\]
		Since $\mathcal{E}$ and $\mathcal{M}$ are closed under composition we know that $e_h\circ e_f\in \mathcal{E}$ and $m_g \circ m_h\in \mathcal{M}$, thus  these arrows gives a $(\mathcal{E}, \mathcal{M})$-factorization of $g\circ f$. On the other hand $g\circ f\in \mathcal{E}$ ($g\circ f\in \mathcal{M}$), thus point $1$ above  implies that  $e_h\circ e_f$ ($m_g\circ m_h$) is an isomorphism. In particular:
		\[(e_h\circ e_f)^{-1}\circ e_h \circ e_f =\id{X} \qquad \text{(}m_g\circ m_h \circ (m_g\circ m_h)^{-1}=\id{Z}\text{)}  \]
		so $e_f$ has a retraction ($m_g$ has a section). The thesis now follows since $e_f$ is epic ($m_g$ is mono). \qedhere 
	\end{enumerate}
\end{proof}

\begin{definition} \index{source} \index{sink} \index{pushout!wide -} \index{pullback!wide -}
	Given a set $I$, a  \emph{source}  (\emph{sink}) is a family $\{f_i\}_{i\in I}$ of arrows  $f_i\colon X\to Y_i$ ($f_i\colon Y_i\to X$) with the same (co)domain. A $\emph{wide pushout}$ (pullback) is the colimit (limit) of a source (sink). We will use $c_i$ ($p_i$) to denote the coprojection from $Y_i$ (the projection to $Y_i$) and $c_X$ to denote the one from $X$.
\end{definition}

\begin{remark}\label{rem:copro}
	Given a wide pushout $(C, \{c_i\}_{i\in I\cup \{X\}})$ on a source $\{f_i\}_{i\in I}$ with $f_i\colon X\to Y_i$, the coprojection $c_X$ is such that, for every $i\in I$, the following diagram commute
	\[\xymatrix{&X\ar[dl]_{f_i} \ar[dr]^{c_X}\\
		Y_i\ar[rr]_{c_{i}} && C}\]
\end{remark}

\begin{proposition} \label{prop:mp} For every proper factorization system $(\mathcal{E}, \mathcal{M})$ on a category $\X$ the following hold:
	\begin{enumerate}
		\item  for every pushout square as the one below,  $e\in \mathcal{E}$ implies $n\in \mathcal{E}$ 
		\[\xymatrix{X\ar[r]^{g} \ar[d]_{e} & V \ar[d]^{n} \\ Y \ar[r]_{f}  & Z}\] 
		\item  if $(C, \{c_i\}_{i\in I\cup \{X\}})$ is a wide pushout on a source $\{e_i\}_{i\in I}$ such that $e_i\colon X\to Y_i$ is in $\mathcal{E}$ for every $i\in I$, then every coprojection is in $\mathcal{E}$ too. 
	\end{enumerate}
\end{proposition}
\begin{proof}
	\begin{enumerate}
		\item 
		Take a pushout square
		\[\xymatrix{X\ar[r]^{g} \ar[d]_{e} & V \ar[d]^{n} \\ Y \ar[r]_{f}  & Z}\] 
		with $e$ in $\mathcal{E}$. By hypothesis  $n=m_n\circ e_n$ for $m_n\colon E\to Y$ in $ \mathcal{M}$  and $e_n\colon V\to E$ in  $\mathcal{E}$.  If we show that $m_n$ is an isomorphism we are done. We can apply again the left lifting property to get $l\colon E\to X$ which makes the following diagram commute.
		\[\xymatrix{X \ar[d]_{e}\ar[r]^{g}&  V \ar[r]^{e_n} &E\ar[d]^{m_n}\\  Z  \ar[urr]^{l} \ar[rr]_{f} && Y}\]
		Therefore we get another diagram
		\[\xymatrix{ X \ar[d]_{e}\ar[r]^{g}& V \ar@/^.5cm/[ddr]^{e_n} \ar[d]^{n} \\
			Z\ar@/_.5cm/[drr]_{l} \ar[r]_{f} & Y \ar@{.>}[dr]^{k}\\
			& &E }\]
		and thus we can deduce the existence of the dotted $k\colon E\to V$. On the one hand, computing we get
		\[\begin{split}
			m_n\circ k\circ n &=m_n  \circ  e_n\\ &=n
		\end{split}\qquad \begin{split}
			m_n\circ k\circ f &=m_n\circ l\\ &=f \end{split}\]
		and so $m_n\circ k= \id{Y}$.  On the other hand 
		\[\begin{split}
			m_n\circ k \circ m_n &= \id{Y} \circ m_n \\&= m_n
		\end{split} \qquad 
		\begin{split}
			k \circ m_n \circ e_n &= k\circ n \\&= e_n
		\end{split}
		\]	
		so the following diagaram commutes 
		\[
		\xymatrix@C=50pt{V \ar[d]_{e_n}\ar[r]^{e_n}& E \ar[d]^{m_n} \\
			E\ar[ur]^{k\circ m_n}\ar[r]_{m_n} & Y}\]
		and thus  $k\circ m_n=\id{E}$ by the uniqueness clause of the left lifting property. Therefore $e_n$ is an isomorphism and the thesis now follows from  point $1$ of \cref{cor:iso}.
		\item By \cref{rem:copro} and point three of \cref{cor:iso} it is enough to show that $c_X$ is in $\mathcal{E}$. Since $(\mathcal{E}, \mathcal{M})$ is a factorization system then $c_X=m\circ e$ for some $m\colon E\to C$ in $\mathcal{M}$ and $e\colon X\to E$ in $\mathcal{E}$ to get, from \cref{rem:copro}, a square
		\[\xymatrix{X\ar[r]^{e} \ar[d]_{e_i} & V \ar[d]^{m} \\ Y_i \ar@{.>}^{k_i}[ur] \ar[r]_{c_i}  & Z}\]
		and the left lifting property provides, for every $i\in I$, the dotted arrow $k_i\colon Y_i\to e$. Let $k$ be the the induced arrow $Z\to V$. Then
		\begin{align*}
			m\circ k\circ c_i&=m\circ k_i\\&=c_i
		\end{align*}
		hence $m\circ k=\id{Z}$. By \cref{cor:iso} $m\in \mathcal{M}$ thus it is an isomorphism and we can conclude. \qedhere 
	\end{enumerate}
\end{proof}

@@

NOTES ON THE FUNDAMENTAL LEMMA WE NEED\todo{da integrare sopra}

@@

\todo{integrare sopra}
\newpage
QUOZIENTI VARI

The setting from now on is: $\X$ $\mathcal{M}$-adhesive, $(\mathcal{E}, \mathcal{M})$ a proper (maybe also stable) factorization system. $\X$ has all pushouts.

\begin{lemma}\label{lem:pushandpullpb}
	Let $\X$ be a category and suppose that the diagram below is given.
	\[\xymatrix{ P \ar[r]^{p_0} \ar[d]_{p_1}& A  \ar@/^.2cm/[ddr]^{k_0}\ar[d]^{f}\\  B  \ar@/_.2cm/[drr]_{k_1} \ar[r]_g & C \ar[dr]^{h}\\&& D }\]
	Then the following hold true:
	\begin{enumerate}
		\item if the outer boundary is a pullback, then the inner square is a pullback too;
		\item if the inner square is a pullback and $h$ is mono then also the outer square is a pullback.
	\end{enumerate}
\end{lemma}
\begin{proof}\begin{enumerate}
		\item Let $e_0\colon E\to A$ and $e_1\colon E\to B$ be a pair of arrows such that 
		\[f\circ e_0=g\circ e_1\]
		Then we also have 
		\begin{align*}
		k_0\circ e_0 &= h\circ f\circ e_0\\&=h\circ g \circ e_1\\&=k_1\circ e_1
		\end{align*}
		Thus there exists a unique arrow $e\colon E\to P$ such that
		\[p_0\circ e= e_0 \qquad p_1\circ e=e_1\]
		which entails our claim.
		\item We proceed as before considering two arrows $e_0\colon E\to A$ and $e_1\colon E\to B$, but in this case we suppose that 
		\[k_0\circ e_0=k_1\circ e_1\]
		In particular, we have
		\begin{align*}
			h\circ f \circ e_0 &= k_0\circ e_0\\&=k_1\circ e_1\\&=h\circ g\circ e_1
		\end{align*}
		Thus, since $h$ is mono, we can deduce that $f\circ e_0$ and $g\circ e_1$ are equal and the thesis now follows immediately.
		\qedhere 
	\end{enumerate}
	\end{proof} 
	In particular, we will use the first point of the previous lemma to deduce the following result.

	\begin{corollary}\label{cor:cond} Let $f\colon A\to C$, $g\colon B\to C$ and $h\colon C\to D$ be arrows in a category $\X$, let also $l_0\colon A'\to A$, $l_1\colon B'\to B$ be two monomorphisms. Suppose moreover, that the following squares are pullbacks:
		\[\xymatrix@C=35pt{Q \ar[r]^{q_0} \ar[d]_{q_1} & A' \ar[d]^{h\circ f\circ l_0} & P \ar[r]^{p_0} \ar[d]_{p_1} & A \ar[d]^{f} & Q  \ar[r]^{l_0\circ q_0} \ar[d]_{l_1\circ q_1}& A \ar[d]^{h\circ f}\\ B \ar[r]_{h\circ g\circ l_1} &D& B \ar[r]_{g}& C &  B \ar[r]_{h\circ g} & D}\]
		Then there exists an arrow $\phi\colon P\to Q$ such that $(P, \{q_i\circ \phi\}_{i=0}^1)$ is a pullback for $f\circ l_0$ along $g\circ l_1$.
	\end{corollary} 
	\begin{proof}
To prove the existence of $\phi\colon P\to Q$ it's enough to consider the solid part of the diagram below.
\[\xymatrix@C=30pt{P \ar@{.>}[dr]^{\phi} \ar@/^.2cm/[drr]^{p_0} \ar@/_.2cm/[ddr]_{p_1}\\ & Q  \ar[r]^(.4){l_0\circ q_0} \ar[d]^{l_1\circ q_1}& A \ar[d]^{h\circ f} \\ &B \ar[r]_{h\circ g} & D}\]
Notice that $\phi$ is a monomorphism. If $\phi\circ a=\phi \circ b$ for some $a,b\colon T\rightrightarrows P$ then
\[\begin{split}
	p_0\circ a&=l_0\circ q_0\circ \phi \circ a\\&=l_0\circ q_0\circ \phi \circ b\\&=p_0\circ b
\end{split} \qquad \begin{split}
p_1\circ a&=l_1\circ q_1\circ \phi \circ a\\&=l_1\circ q_1\circ \phi \circ b\\&=p_1\circ b
\end{split} \]

Moreover, we also have the following chain of equalities, showing that  $(P, \{q_i\circ \phi\}_{i=0}^1)$  is a candidate to be a pullback of $f\circ l_0$ along $g\circ l_1$.
\begin{align*}
f\circ l_0\circ q_0\circ \phi &=f\circ p_0\\&=g\circ p_1\\&=g\circ l_1\circ q_1\circ \phi 
\end{align*}
Let now $e_0\colon E\to A'$ and $e_1\colon E\to B'$ be arrows such that 
\[f\circ l_0\circ e_0=g\circ l_1\circ e_1\]
By hypothesis $(P, \{p_i\}_{i=0}^1)$ is a pullback of $f$ along $g$, therefore we there is a unique  arrow $e\colon E\to P$ such that
\[p_0\circ e = l_0\circ e_0 \qquad p_1\circ e=l_1\circ e_1\]
But then the two identities
\[\begin{split}
	l_0\circ q_0\circ \phi\circ e&=p_0\circ e\\& =l_0\circ e_0 \end{split} \qquad \begin{split}l_1\circ q_1\circ \phi\circ e&=p_1\circ e\\& =l_1\circ e_1
\end{split}\]
allow us to conclude that 
\[q_0\circ \phi\circ e=l_0\circ e_0 \qquad q_1\circ \phi\circ e=l_1\circ e_1\]

Finally, if $e'\colon E \to P$ is another arrow such that 		
\[q_0\circ \phi\circ e'=l_0\circ e_0 \qquad q_1\circ \phi\circ e'=l_1\circ e_1\]
then we have 
\[q_0\circ \phi\circ e'=q_0\circ \phi\circ e \qquad q_1\circ \phi\circ e'=q_1\circ \phi\circ e\]
Therefore $\phi\circ e'=\phi \circ e$ and we can conclude since $\phi$ is mono.
	\end{proof}

\begin{lemma}
	Da ind pair a pb
\end{lemma}

\begin{proof}
	contenuto...
\end{proof}

\begin{lemma}
	da pb a ind pair
\end{lemma}
\begin{proof}
	contenuto...
\end{proof}

\begin{corollary}
	caratterizzazione
\end{corollary}

@@

\section{$\mathcal{M}$-adhesivity is not enough}\label{app:fill}
 In \Cref{prop:equi} we proved that, in the linear case, the existence of an indpendence pair between two derivation is equivalent to that of a filler between them. This result can be further refined: in a\cite{baldan2011adhesivity} a class  $\mathbb{B}$ of (quasi)adhesive category is defined for which the local Church-Rosser Theorem holds even for left-linear DPO-rewriting system. In our language, and given \Cref{prop:fil} and \Cref{rem:locCR}, this amount to prove that, for elements of $\mathbb{B}$, every independence pair induces a filler.
